# Matrices in R
**Created by John Palowitch, UNC Chapel Hill**

**Edited by Jan Hannig, UNC Chapel Hill**


### Conceptual Questions 
Please read the document ***Matrices in R.docx*** before completing this assignment. This assignment has two goals:

-Learning new commands in R from a written source.

-Using R for Matrix Calculations
<br><br>
1. Suppose:
\[
\mathbf{A} = \begin{bmatrix}2&3&4&5\\-1&0&3&2\end{bmatrix},\;\;\;\;\mathbf{B} = \begin{bmatrix}4&-1&10&0\\-2&-2&5&6\end{bmatrix}
\]
\[
\mathbf{a} = \begin{bmatrix}6&2&-11&1\end{bmatrix},\;\;\;\;\mathbf{b} = \begin{bmatrix}4\\2\\3\\0\end{bmatrix},\;\;\;\;\mathbf{c} = \begin{bmatrix}-3\\-10\end{bmatrix}
\]
a. Write down the dimensions of the following (you can use a comment box that you create). If the operation is not possible, write "N.P."

i.$\mathbf{A}^T\mathbf{B}$   
```
4X4
```
ii.$\mathbf{B}\mathbf{A}^T$  
```
2X2 
```
iii.$\mathbf{A}\mathbf{B}$   
```
N.P. The ncol(A) = 4 is not equal to the nrow(B) = 2.
```
iv.$\mathbf{A}^T\mathbf{A}$  
```
4X4
```
v.$\mathbf{b}^T\mathbf{A}$   
```
N.P. The ncol(t(b)) = 4 is not equal to the nrow(A) = 2.
```
vi.$\mathbf{a}\mathbf{A}^T$  
```
1X2
```
vii.$\mathbf{B}\mathbf{b}$   
```
2X1
```
viii.$\mathbf{B}\mathbf{c}^T$ 
```
N.P. The ncol(B) = 4 is not equal to the nrow(t(c)) = 1.
```
ix.$\mathbf{a}\mathbf{c}$    
```
N.P. The ncol(a) = 4 is not equal to the nrow(c) = 2.
```
x.$\mathbf{b}\mathbf{a}$     
```
4X4
```

b. Again, write down the dimensions of the following (on this sheet, next to the expression). If the operation is not possible, write "N.P."

i.$(\mathbf{A}^T\mathbf{B})^{-1}\mathbf{a}$
```
N.P. Matrix t(A) %*% B is not invertible.
```
ii.$\mathbf{c}(\mathbf{B}\mathbf{A}^T)^{-1}$
```
N.P. The ncol(c) = 1 is not equal to nrow(solve(B %*% t(A))) = 2.
```
iii.$(\mathbf{B}^T\mathbf{B})^{-1}\mathbf{B}^T\mathbf{c}$
```
N.P. Matrix t(B) %*% B is singular.
```
iv.$\mathbf{b}^T\mathbf{A}^T\mathbf{c}$
```
1X1
```
v.$(\mathbf{B}\mathbf{B}^T)^{-1}\mathbf{B}\mathbf{b}$
```
2X1
```     
        
### Applied Questions 
2. In a code block below, create and print the following matrices and vectors:
\[
\mathbf{A} = \begin{bmatrix}-6&0&1\\1&10&7.5\end{bmatrix},\;\;\;\;\mathbf{B} = \begin{bmatrix}-15&1&1\\5.25&1&2.5\end{bmatrix}
\]
\[
\mathbf{a} = \begin{bmatrix}6&2&1\end{bmatrix},\;\;\;\;\mathbf{b} = \begin{bmatrix}2\\-7\end{bmatrix}
\]
```{r, eval = TRUE}
A <- matrix(c(-6,0,1, 1,10,7.5), nrow = 2, ncol = 3, byrow = TRUE)
print(A)
B <- matrix(c(-15,1,1, 5.25,1,2.5), nrow = 2, ncol = 3, byrow = TRUE)
print(B)
a <- matrix(c(6,2,1), nrow = 1, ncol = 3, byrow = TRUE)
print(a)
b <- matrix(c(2,-7), nrow = 2, ncol = 1, byrow = FALSE)
print(b)
```

3. Compute (with **R**, in a new code block below) and print the following products:
  a. $\mathbf{A}^T\mathbf{B}$
```{r, eval = TRUE}
A <- matrix(c(-6,0,1, 1,10,7.5), nrow = 2, ncol = 3, byrow = TRUE)
B <- matrix(c(-15,1,1, 5.25,1,2.5), nrow = 2, ncol = 3, byrow = TRUE)
print(t(A)%*%B)
```
  b. $\mathbf{B}\mathbf{a}^T$
```{r, eval = TRUE}
B <- matrix(c(-15,1,1, 5.25,1,2.5), nrow = 2, ncol = 3, byrow = TRUE)
a <- matrix(c(6,2,1), nrow = 1, ncol = 3, byrow = TRUE)
print(B%*%t(a))
```
  c. $\mathbf{b}^T\mathbf{A}$
```{r, eval = TRUE}
b <- matrix(c(2,-7), nrow = 2, ncol = 1, byrow = FALSE)
A <- matrix(c(-6,0,1, 1,10,7.5), nrow = 2, ncol = 3, byrow = TRUE)
print(t(b)%*%A)
```
  d. $\mathbf{A}^T\mathbf{b}$
```{r, eval = TRUE}
A <- matrix(c(-6,0,1, 1,10,7.5), nrow = 2, ncol = 3, byrow = TRUE)
b <- matrix(c(2,-7), nrow = 2, ncol = 1, byrow = FALSE)
print(t(A)%*%b)
```
  e. $\mathbf{a}^T\mathbf{a}$
```{r, eval = TRUE}
a <- matrix(c(6,2,1), nrow = 1, ncol = 3, byrow = TRUE)
print(t(a)%*%a)
```
  f. $\mathbf{a}\mathbf{a}^T$
```{r, eval = TRUE}
a <- matrix(c(6,2,1), nrow = 1, ncol = 3, byrow = TRUE)
print(a%*%t(a))
```
  g. $(\mathbf{A}^T\mathbf{A})^{-1}$
```
Matrix t(A)%*%A is computationally singular, so solve(t(A)%*%A) does not exist.
```
  h. $(\mathbf{A}^T\mathbf{A})^{-1}\mathbf{A}^T\mathbf{b}$
```
Matrix t(A)%*%A is computationally singular, so solve(t(A)%*%A)%*%t(A)%*%b does not exist.
```
  i. $\mathbf{b}^T\mathbf{A}\mathbf{a}^T$
```{r, eval = TRUE}
b <- matrix(c(2,-7), nrow = 2, ncol = 1, byrow = FALSE)
A <- matrix(c(-6,0,1, 1,10,7.5), nrow = 2, ncol = 3, byrow = TRUE)
a <- matrix(c(6,2,1), nrow = 1, ncol = 3, byrow = TRUE)
print(t(b)%*%A%*%t(a))
```
  j. $\mathbf{a}\mathbf{B}\mathbf{A}^T\mathbf{b}$
```
The ncol(a) = 3 is not equal to the nrow(B) = 2, so a%*%B%*%t(A)%*%b is not valid.
```


